Solving the Poisson-equation in one dimension \\ \hrulefill\small{ FYS3150: Computational Physics }\hrulefill

### 2016-09-16 Project1/article/article.tex
Conclusion

By looking at the results one is led to believe that there are more reasons to prefer the tridiagonal matrix algorithm for this kind of problem. When using the LU-decomposition method the short term memory is quickly filled with nothing but the matrix $A$. Choosing $n=10^5$ is equivalent to $8\times10^{10}$ bytes$=80$GB of memory! A brand new computer ships with around $8$ GB of memory today, which is a tenth of the space required. Obviously, this inhibits our ability to perform the calculation.

\section{Conclusion}

\begin{thebibliography}{10}
  \bibitem{thomasalgo}{Thomas, L.H. (1949), \emph{Elliptic
        Problems in Linear Differential Equations over a
        Network}. Watson Sci. Comput. Lab Report, Columbia University,
      New York.}
    \bibitem{morten}{Hjorth-Jensen, M. (2016). \emph{Computational
        Physics - Lecture Notes 2016}. University of Oslo}
    \bibitem{golub}{Golub, G.H., van Loan, C.F. (1996). \emph{Matrix
          Computations} (3rd ed.), Baltimore: John Hopkins.}
\end{thebibliography}

>>>>

By looking at the results one is led to believe that there are more reasons to prefer the tridiagonal matrix algorithm for this kind of problem. When using the LU decomposition method the short term memory is quickly filled with nothing but the matrix $A$. Choosing $n=10^5$ is equivalent to $8\times10^{10}$ bytes$=80$GB of memory! A brand new computer ships with around $8$ GB of memory today, which is a tenth of the space required. Obviously, this inhibits our ability to perform the calculation.

\section{Conclusion}
As we have seen, there is significant advantage to fine-tuning and picking the correct algorithm before solving a problem numerically. We have experienced the restrictions in memory when computing, and have been made aware of the problem this poses when doing numerical calculations. The specialised tridiagonal matrix algorithm was able to solve the one-dimensional Poisson equation more than $20000$ times faster than the more general LU decomposition method. This can be explained by the fact that the LU decomposition method require $\mathcal{O}(\frac{2}{3}n^3)$ FLOPS while the specialized tridiagonal matrix algorithm will do with a meager $\mathcal{O}(8n)$ FLOPS.

\begin{thebibliography}{10}
  \bibitem{thomasalgo}{Thomas, L.H. (1949), \emph{Elliptic
        Problems in Linear Differential Equations over a
        Network}. Watson Sci. Comput. Lab Report, Columbia University,
      New York.}
    \bibitem{morten}{Hjorth-Jensen, M. (2015). \emph{Computational
        Physics - Lecture Notes 2015}. University of Oslo}
    \bibitem{golub}{Golub, G.H., van Loan, C.F. (1996). \emph{Matrix
          Computations} (3rd ed.), Baltimore: John Hopkins.}
\end{thebibliography}
